---
title: Simulating an interactive Twitter graph
---
```{r, include=F}
opts_chunk$set(echo=F, message=F, warning=F, error=F)
source('project.R')
library(plotly)
py <- plotly()
```

This week I submitted some early work to a workshop co-located with EuroSys in Bordeaux, France. In the process of gathering data for my submission, I discovered some interesting things about simulating social network graphs that I thought I'd share more broadly.

## Followers following power law
Everyone knows the degree distribution of natural graphs follows a power law: most people have a handful of friends or followers (in my case, on the order of 0), while a handful of people (Katy Perry, Justin Bieber, Barack Obama) have many millions of followers. We know how to at least reproduce graphs with this same degree distribution with a class of synthetic graph generators known as Kronecker generators.

```{r, include=F}
# library(jsonlite)
# df.histogram <- function(json, version="none") {
#   d <- fromJSON(json)
#   return(data.frame(x=num(names(d)), y=num(unlist(d)), version=version))
# }
# df <- subset(db("
#     select * from tapir where stat_following_counts is not null
#     and name like '%v0.14%'
# "),
#   nclients == 32
#   & initusers == 4096
# )
# df$grp <- with(df, sprintf("%s\n%s\nmix:%s/%s,\n%s", name, ccmode, mix, alpha, gen))
# 
# histogram.facets <- function(df, measure, grp) {
#   d <- data.frame(x=c(),y=c(),version=c())
#   for (i in 1:nrow(df)) {
#     d <- rbind(d, df.histogram(df[i,measure], df[i,grp]))
#   }
#   return(d)
# }
```

```{r followers, include=F}
# d.follow <- histogram.facets(subset(df,
#     initusers == 4096 & mix == 'geom_repost'
# ), 'stat_follower_counts', 'grp')
# ggplot(d.follow, aes(x=x, weight=y))+
#     stat_ecdf(color=c.blue)+
#     xlab('# followers / user (log scale)')+ylab('CDF (log scale)')+
#     scale_x_log10(breaks=c(1,10,100,1000))+scale_y_log10(breaks=c(0.1,0.5,1.0))+
#     theme_mine
# py$ggplotly()
```

```{r throughput, plotly=TRUE}
d <- data(db("
    select * from tapir where 
    generator_time is not null and total_time is not null
    and name like 'claret-v0.14%'
  ",
    factors=c('nshards', 'nclients'),
    numeric=c('total_time', 'txn_count')
))

d.u <- subset(d, nshards == 4 & initusers == 4096 
                & grepl('geom_repost|read_heavy', mix))
                
                

g <- ggplot(d.u, aes(x=nclients, y=throughput/1000, group=cc, fill=cc, color=cc, linetype=cc))+
    stat_summary(fun.y=mean, geom="line")+
    xlab('Concurrent clients')+ylab('Throughput (k trans. / sec)')+
    # expand_limits(y=0)
    scale_y_continuous(limits=c(0, 25))+
    facet_wrap(~workload)+
    theme_mine+
    cc_scales(title='Concurrency\ncontrol:')

py$ggplotly(g, session="knitr")
```

